### 2020-03-11 scripts/00_data_wrangling.Rmd
subsidy map back up

---
output:
  html_document: default

title: "SubsidyExplorer - Step 0: Data Wrangling"
author: "Kat Millage, Vienna Saccomanno, Laura Lea Rubino, Christopher Costello"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
---

```{r setup, include=FALSE}
library(knitr) 
library(countrycode) # country name matching
library(WDI) # World Bank GDP and population numbers by country
library(janitor) # cleaning column names
library(tidyverse)
library(readxl)
```

```{r}
# Code options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = F)

# Directories for the output files generated by this script
data_dir <- here::here("data", "edited/")
if (dir.exists(data_dir) == F) {
  dir.create(data_dir, recursive = T)
}

app_data_dir <- here::here("app", "data/")
if (dir.exists(app_data_dir) == F) {
  dir.create(app_data_dir, recursive = T)
}

results_dir <- here::here("results", "00-data-wrangling/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}

```

# Introduction

This script handles the preliminary data wrangling for the raw datasets underlying the SubsidyExplorer toolkit. It outputs tidyed data files to `r data_dir`.

It also creates and outputs four aggregated CSV files that are used as input data by the SubsidyExplorer toolkit (saved to `r app_data_dir`): 
- subsidies-2009-2018-tidy.csv
- demographic-2000-2018-tidy.csv
- fisheries-2000-2017.csv
- fao-landings-group-2000-2017.csv

Summary figures are exported to `r results_dir`.

```{r}
# Paths to subsidies data files
sumaila_subsidies_path <- here::here("data", "raw", "Sumaila-et-al-2019-subsidies", "Subsidies_PEW_Final.xlsx")
oecd_fse_path <- here::here("data", "raw", "OECD-2019-fse", "FISH_FSE_10032020183349581.csv")

# Paths to demographic data files
fao_fishers_path <- here::here("data", "raw", "FAO-2019-yearbook", "fao_fishers_yearbook_070619.csv")
teh_fte_path <- here::here("data", "raw", "Teh-and-Sumaila-2013-fte", "Fisher employment FTE.xlsx")

# Paths to fisheries data files
fao_production_path <- here::here("data", "raw", "FAO-2019-capture-production", "TS_FI_CAPTURE.csv")
#fao_value_path
melnychuk_price_path <- here::here("data", "raw", "Melnychuk-et-al-2013-price", "Melnychuk_Exvessel_Price_Database.csv")

# Paths to misc data files 
costello_upsides_path <- here::here("data", "raw", "Costello-et-al-2016-upsides", "ProjectionData_2012.csv")
iuu_vessel_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")
ram_sadb_path <- here::here("data", "raw", "RAM-2018-stock-assessment-database", "RLSADB v4.40", "DB Files With Assessment Data", "RLSADB v4.40 (assessment data only).xlsx")
```

The following data sources are included in this analysis:

Estimates of annual fisheries subsidies provision by state from Sumaila et al. (2019) in 2018 US\$. These data were provided to us directly by the authors as a .xlsx file and are stored at `r sumaila_subsidies_path`.
> Sumaila, U. R., Ebrahim, N., Schuhbauer, A., Skerritt, D., Li, Y., Kim, H. S., … Pauly, D. (2019). Updated estimates and analysis of global fisheries subsidies. Marine Policy, 109, 103695. https://doi.org/10.1016/j.marpol.2019.103695

Estimates of fisheries support provision by state from the OECD (2008 - 2018). These data were last downloaded from the OECD iLibrary portal as a CSV file on March 10, 2019 and are stored as `r oecd_fse_path`.
> OECD (2019), "Fisheries: Fisheries support estimates", OECD Agriculture Statistics (database), https://doi.org/10.1787/ade64fdc-en (accessed on 10 March 2019).

Total GDP and population by state from the World Bank's World Development Indicators Database (2000 - 2018). These data were accessed through the `WDI` package. The code used to pull the data is below. 
> The World Bank (2019). "World Development Indicators", https://datacatalog.worldbank.org/dataset/world-development-indicators (accessed on 24 October 2019).

Total number of fishers by state from the FAO Yearbook of Fishery and Aquaculture Statistics (misc. years between 1995 and 2017). These data were manually transcribed from PDF (`./data/raw/fleet_&_employment.pdf`) to CSV format on July 6, 2019 and are stored as `r fao_fishers_path`. 
> FAO (2019). FAO yearbook. Fishery and Aquaculture Statistics 2017, http://www.fao.org/fishery/statistics/global-fishers/en (accessed on 24 October 2019).

Estimated number of full-time jobs in fisheries from Teh and Sumaila (2013) in 2003. These data were provided to us directly by the authors as a .xlsx file and are stored as `r teh_fte_path`. 
> Teh, L. C. L., & Sumaila, U. R. (2011). Contribution of marine fisheries to worldwide employment. Fish and Fisheries, 14(1), 77–88. https://doi.org/10.1111/j.1467-2979.2011.00450.x

Fisheries landings in 2017 from the FAO Global Capture Production database (2019) and estimates of landed value generated using the ex-vessel price database created by Melnychuk et al. (2017). These data were were wrangled in the previous script, and the edited data are stored as `r catch_dat_path`. 
> FAO (2019). FAO yearbook. Fishery and Aquaculture Statistics 2017, http://www.fao.org/fishery/statistics/global-capture-production/en (accessed on 24 October 2019). 
> Melnychuk, M. C., Clavelle, T., Owashi, B., & Strauss, K. (2017). Reconstruction of global ex-vessel prices of fished species. ICES Journal of Marine Science, 74(1), 121–133. https://doi.org/10.1093/icesjms/fsw169

# Note: Political entities, dependencies, and naming

The naming of flag and coastal states varies widely across our data sources and we recognize that this can be a politially sensitive matter. For the purposes of merging multiple data sources, we therefore translate all state names from the raw data sources to ISO-3 character codes. Country names are only used for display purposes, and the use of any particular name is not meant to convey any options regarding sovereignty on behalf of the authors.

```{r}
# Load country dependencies helper file
country_dependencies <- read_csv(here::here("data", "helper-files", "country_dependencies.csv"))

# EU countries
eu_countries <- country_dependencies$country_iso3[country_dependencies$is_EU]
eu_country_names <- sort(countrycode(eu_countries, "iso3c", "country.name"))

# EU overseas territories
eu_territories <- country_dependencies$country_iso3[country_dependencies$sovereign_iso3 %in% eu_countries & !country_dependencies$is_EU]
eu_territory_names <- sort(c(countrycode(eu_territories, "iso3c", "country.name"), "Ascension Island", "Tristan da Cunha"))

# US overseas territories:  American Samoa, Guam, Northern Marianas Islands, Puerto Rico, US Virgin Islands
us_territories <- country_dependencies$country_iso3[country_dependencies$sovereign_iso3 == "USA"]
us_territory_names <- sort(countrycode(us_territories, "iso3c", "country.name"))
```

Additionally, it is important to note that the European Union as a political entity is a Member of the World Trade Organization, rather then the individual states that make up the European Union. We therefore need to create a profile for the European Union as a whole.  The following `r length(eu_country_names)` states are considered to be part of the European Union for the purposes of this analysis: `r paste(eu_country_names, collapse = ", ")`.

The Netherlands, France, Denmark, and the United Kingdom also have overseas dependencies that should be considered to be a part of the European Union for statistical purposes. The following overseas dependencies are also considered to be a part of the European Union for the purpose of aggregating data in this analysis: `r paste(eu_territory_names, collapse = ", ")`.

Outside of the EU, the only other dependencies we need to worry about are those that relate to the United States. The following overseas dependencies are considered to be a part of the United States for the purpose of aggregating data in this analysis: `r paste(us_territory_names, collapse = ", ")`.

# Section 1 - Subsidies Data

## Updated estimates and analysis of global fisheries subsidies from Sumaila et al. (2019)

```{r}
### Data Wrangling
# Load and wrangle 2019 data
sumaila_2019_dat <- read_xlsx(sumaila_subsidies_path) %>%
  clean_names(case = "snake") %>%
  mutate(iso3 = case_when(country == "Dominican Rp" ~ "DOM",
                          country == "Micronesia" ~ "FSM",
                          country == "Untd Arab Em" ~ "ARE",
                          TRUE ~ countrycode(country, "country.name", "iso3c"))) %>%
  dplyr::select(iso3, type = class, data_type, value = constant_2018_usd) %>%
  mutate(year = 2018,
         units = "2018 US$",
         source = "Sumaila et al. (2019)") %>%
  arrange(iso3)

# Load lookup table for classification scheme
sumaila_lookup <- read_csv(here::here("data", "helper-files", "sumaila_subsidies_classification.csv"))

# Clean up the naming/spelling of the classification scheme as well as data type
sumaila_2019_dat_edit <- sumaila_2019_dat %>%
  left_join(sumaila_lookup, by = "type") %>%
  dplyr::select(iso3, year, category, type, type_name = subtype, value, data_type, units, source)

sumaila_2019_dat_edit$data_type[sumaila_2019_dat_edit$data_type == "not found evidence of subsidy"] <- "Found no evidience of subsidy"

# Deal with the EU
sumaila_2019_dat_eu <- sumaila_2019_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU",
         data_type = "Aggregate") %>%
  group_by(entity, year, category, type, type_name, data_type, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity)

# Add the EU back in
sumaila_2019_dat_edit <- sumaila_2019_dat_edit %>%
  bind_rows(sumaila_2019_dat_eu)

# Save
write_csv(sumaila_2019_dat_edit, paste0(data_dir, "sumaila_et_al_2019_subsidies_tidy.csv"))
write_csv(sumaila_2019_dat_edit, paste0(app_data_dir, "sumaila_et_al_2019_subsidies_tidy.csv"))

### Summary Plots

```

## Estimates of fisheries support provision by state from the OECD (2008 - 2018)

```{r}
# Load and wrangle OECD data (2009 - 2017)
oecd_fse_dat <- read_csv(oecd_fse_path) %>%
  clean_names(case = "snake") %>%
  dplyr::select(iso3 = country, country = country_2, code = variable, name = variable_2, measure, year, units = unit_code, value) %>%
  dplyr::filter(measure == "USD", # only pull estimates in USD for consistancy
                iso3 != "OECD") # remove totals

# Annoying, but the OECD classification scheme is a little more complicated and the type codes don't tell us anything about the classification scheme just by looking at them so we are going to manually assign more meaningful ones to make sure we're not double counting. 
oecd_types_table <- oecd_dat %>%
  distinct(code, name)

# Create blank lookup table
oecd_lookup <- tibble(category = character(0),
                                  category_name = character(0))

# Categories
categories <- enframe(
  c("0" = "TRANSFERS TO INDIVIDUAL FISHERS - Non Budgetary",
    "1" = "TRANSFERS TO INDIVIDUAL FISHERS - Budgetary",
    "2" = "GENERAL SERVICE SUPPORT ESTIMATE",
    "3" = "COST RECOVERY CHARGES"), name = "category", value = "category_name")

oecd_lookup <- oecd_lookup %>%
  bind_rows(categories)

# Non budgetary transfers to individual fishers
types <- enframe(
  c("0A" = "Market price support",
    "0B" = "Fuel tax concessions",
    "1A" = "Transfers based on input use",
    "1B" = "Transfers based on fishers income",
    "1C" = "Transfers based on the reduction of productive capacity",
    "1D" = "Miscellaneous transfers to fishers",
    "2A" = "Access to other countries’ waters",
    "2B" = "Provision of infrastructure",
    "2C" = "Marketing and promotion",
    "2D" = "Support to fishing communities",
    "2E" = "Education and training",
    "2F" = "Research and development",
    "2G" = "Management of resources",
    "2H" = "Miscellaneous transfers to general services",
    "3A" = "Cost Recovery Charges, for resource access rights",
    "3B" = "Cost Recovery Charges, for infrastructure access",
    "3C" = "Cost Recovery Charges, for management, research and enforcement",
    #"3-D" = "Resource rent taxes and charges", #doesn't appear in this data set?
    "3E" = "Cost Recovery Charges, other"), name = "type", value = "type_name") %>%
  mutate(category = str_replace(type, "[A-Z]", ""))

oecd_lookup <- oecd_lookup %>%
  left_join(types, by = "category")

# Subtypes
subtypes <- enframe(
  c("1A-1" = "Transfers based on variable input use",
    "1A-2" = "Transfers based on fixed capital formation",
    "1B-1" = "Income support",
    "1B-2" = "Special insurance system for fishers",
    "2B-1" = "Capital expenditures",
    "2B-2" = "Subsidized access to infrastructure",
    "2G-1" = "Management expenditures",
    "2G-2" = "Stock enhancement programs",
    "2G-3" = "Enforcement expenditures"), name = "subtype", value = "subtype_name") %>%
  mutate(type = str_replace(subtype, "-[^-]*$", ""))

oecd_lookup <- oecd_lookup %>%
  left_join(subtypes, by = "type")

# Sub-subtypes
ssubtypes <- enframe(
  c("1A-2a" = "Support to vessel construction/purchase",
    "1A-2b" = "Support to modernisation",
    "1A-2c" = "Support to other fixed costs"), name = "ssubtype", "ssubtype_name") %>%
  mutate(subtype = str_replace(ssubtype, "[a-z]$", ""))

oecd_lookup <- oecd_lookup %>%
  left_join(ssubtypes, by = "subtype")

# Now rejoin to get a better sense of organization
oecd_lookup <- oecd_lookup %>%
  left_join(oecd_types_table, by = c("category_name" = "name")) %>%
  rename(category_code = code) %>%
  left_join(oecd_types_table, by = c("type_name" = "name")) %>%
  rename(type_code = code) %>%
  left_join(oecd_types_table, by = c("subtype_name" = "name")) %>%
  rename(subtype_code = code) %>%
  left_join(oecd_types_table, by = c("ssubtype_name" = "name")) %>%
  rename(ssubtype_code = code) %>%
  dplyr::select(category, category_code, category_name, type, type_code, type_name, subtype, subtype_code, subtype_name, ssubtype, ssubtype_code, ssubtype_name)

write_csv(oecd_lookup, paste0(data_dir, "oecd-subsidies-classification.csv"))

# Create variable for matching 
oecd_matching <- oecd_lookup %>%
  mutate(code_for_matching = case_when(!is.na(ssubtype_code) ~ ssubtype_code,
                                       !is.na(subtype_code) ~ subtype_code,
                                       !is.na(type_code) ~ type_code))

# Add better descriptors back into data set
oecd_dat_edit <- oecd_dat %>%
  dplyr::select(-name) %>%
  inner_join(oecd_matching, by = c("code" = "code_for_matching")) %>%
  mutate(category_keep = case_when(category == "0" ~ "Non Budgetary Transfers",
                                   category == "1" ~ "Budgetary Transfers",
                                   category == "2" ~ "General Service Support",
                                   category == "3" ~ "Cost Recovery Charges"),
         type_keep = case_when(!is.na(ssubtype_code) ~ ssubtype,
                               !is.na(subtype_code) ~ subtype,
                               TRUE ~ type),
         subtype_keep = case_when(!is.na(ssubtype_code) ~ ssubtype_name,
                               !is.na(subtype_code) ~ subtype_name,
                               TRUE ~ type_name)) %>%
  dplyr::select(iso3, year, category = category_keep, type = type_keep, subtype = subtype_keep, value) %>%
  mutate(units = paste0(year, " US$"),
         source = "OECD")

# Deal with the EU
oecd_dat_eu <- oecd_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, category, type, subtype, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, type)

# Add EU back in 
oecd_dat_edit <- oecd_dat_edit %>%
  bind_rows(oecd_dat_eu)

write_csv(oecd_dat_edit, paste0(data_dir, "oecd-subsidies-2009-2017-tidy.csv"))

```