### 2020-06-29 scripts/06_run_model.Rmd

---
output:
  html_document: default

title: "SubsidyExplorer - Step 5: Run Model"
author: "Kat Millage, Vienna Saccomanno, Laura Lea Rubino, Christopher Costello"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
bibliography: "../pew-subsidies.bib"
---

```{r setup, include=FALSE}
# Load packages
library(knitr) 
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This script generates the most ambitous global and regional results obtained by removing all capacity enhancing subsidies from all WTO Members and Observers.

```{r}
# Source model script

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.