### 2020-08-06 scripts/01_data_wrangling.Rmd
finalizing data sources

---
output:
  html_document: default

title: "SubsidyExplorer - Step 1: Data Wrangling"
author: "Kat Millage, Vienna Saccomanno, Laura Lea Rubino, Christopher Costello"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
---

```{r}
# Packages
library(knitr) 
library(countrycode) # country name matching
library(here)
library(WDI) # World Bank GDP and population numbers by country
library(janitor) # cleaning column names
library(tidyverse)
library(readxl)

# Code chunk defaults
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE, include = FALSE)

# Source helper file
source(here::here("scripts", "00_helpers.R"))

# Directories for the output files generated by this script
data_dir <- here::here("data", "edited/")
if (dir.exists(data_dir) == F) {
  dir.create(data_dir, recursive = T)
}

app_data_dir <- here::here("app", "data/")
if (dir.exists(app_data_dir) == F) {
  dir.create(app_data_dir, recursive = T)
}

results_dir <- here::here("results", "01-data-wrangling/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}
```

# Description

This script handles the preliminary data wrangling for the raw datasets underlying the SubsidyExplorer analysis. It outputs tidyed data files to `r data_dir`. Some of these tidied data files are used directly as input data by the SubsidyExplorer toolkit (saved to `r app_data_dir`).

```{r}
### Paths to demographic data files
# WDI data - no path, accessed through package
fao_fishers_path <- here::here("data", "raw", "FAO-2019-yearbook", "fao_fishers_yearbook_070619.csv")
teh_fte_path <- here::here("data", "raw", "Teh-and-Sumaila-2013-fte", "Fisher employment FTE.xlsx")

### Paths to fisheries subsidies data files
sumaila_subsidies_path <- here::here("data", "raw", "Sumaila-et-al-2019-subsidies", "Subsidies_PEW_Final.xlsx")
oecd_fse_path <- here::here("data", "raw", "OECD-2019-fse", "FISH_FSE_10032020183349581.csv")
schuhbauer_subsidies_ssf_path <- here::here("data", "raw", "Schuhbauer-et-al-2017-ssf-subsidies", "Subsidies_SSF_split_2018.csv")

### Paths to other fisheries data files
fao_production_path <- here::here("data", "raw", "FAO-2019-capture-production", "Capture_2019.1.0/")
#fao_value_path
melnychuk_price_path <- here::here("data", "raw", "Melnychuk-et-al-2013-price", "Melnychuk_Exvessel_Price_Database.csv")
ram_sadb_path <- here::here("data", "raw", "RAM-2018-stock-assessment-database", "RLSADB v4.40", "DB Files With Assessment Data", "RLSADB v4.40 (assessment data only).xlsx")
costello_upsides_path <- here::here("data", "raw", "Costello-et-al-2016-upsides", "ProjectionData_2012.csv")
melnychuk_fmi_path <- here::here("data", "raw", "Melnychuk-et-al-2017-fmi", "fmi_values_country.csv")
iuu_vessel_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")

### Paths to other necessary files
cap_tier_path <- here::here("data", "raw", "USA-Cap-Tier-Proposal", "US_cap_tier_data.csv")
country_lookup_path <- here::here("data", "lookup-tables", "country_dependencies.csv")
fao_regions_for_ram_lookup_path <- here::here("data", "lookup-tables", "fao_areas_for_ram_stocks_v4.491.csv")
```

The following data sources are included in this analysis:

**Demographic Data**
Total GDP, GDP from fisheries, forestry, and agriculture, and population by state from the World Bank's World Development Indicators Database (2000 - 2018). These data were accessed through the `WDI` package. The code used to pull the data is below. 
> World Bank. (2020). World Development Indicators. https://data.worldbank.org/

Total number of fishers by state from the FAO Yearbook of Fishery and Aquaculture Statistics (misc. years between 1995 and 2017). These data were manually transcribed from PDF (`./data/raw/fleet_&_employment.pdf`) to CSV format on July 6, 2019 and are stored as `r fao_fishers_path`. 
> FAO. (2019). FAO Yearbook of Fishery and Aquaculture Statistics 2017 (p. 108).

Estimated number of full-time jobs in fisheries from Teh and Sumaila (2013) in 2003. These data were provided to us directly by the authors and are stored as `r teh_fte_path`. 
> Teh, L. C. L., & Sumaila, U. R. (2011). Contribution of marine fisheries to worldwide employment. Fish and Fisheries, 14(1), 77–88. https://doi.org/10.1111/j.1467-2979.2011.00450.x

**Fisheries Subsidies Data**
Estimates of annual fisheries subsidies provision by state from Sumaila et al. (2019) in 2018 US\$. These data were provided to us directly by the authors and are stored at `r sumaila_subsidies_path`.
> Sumaila, U. R., Ebrahim, N., Schuhbauer, A., Skerritt, D., Li, Y., Kim, H. S., … Pauly, D. (2019). Updated estimates and analysis of global fisheries subsidies. Marine Policy, 109, 103695. https://doi.org/10.1016/j.marpol.2019.103695

Estimates of fisheries support provision by state from the OECD (2008 - 2018). These data were last downloaded from the OECD iLibrary portal on March 10, 2019 and are stored as `r oecd_fse_path`.
> OECD. (2019). Fisheries: Fisheries support estimates. Fisheries and Aquaculture Statistics, OECD Agriculture Statistics (Database). https://doi.org/10.1787/ade64fdc-en

Estimates of annual fisheries subsidies provision to small scale fisheries by state in 2018 US\$. Data used in this analysis are unpublished updates (based on the Sumaila et al. (2019) data) from those reported in Schuhbauer et al. (2017) (which are in 2009 US\$). These data were provided to us directly by the authors and are stored at `r schuhbauer_subsidies_ssf_path`. 
> Schuhbauer, A., Chuenpagdee, R., Cheung, W. W. L., Greer, K., & Sumaila, U. R. (2017). How subsidies affect the economic viability of small-scale fisheries. Marine Policy, 82, 114–121. https://doi.org/10.1016/j.marpol.2017.05.013

**Other Fisheries Data**
Fisheries landings in 2018 from the FAO Global Capture Production database (2020). These data were last downloaded from the FAO website on July 25, 2020 and are stored as `r fao_production_path`. 
> FAO. (2020). Global Capture Production Dataset. http://www.fao.org/fishery/statistics/global-capture-production/en

Estimates of ex-vessel prices created by Melnychuk et al. (2017). These data were provided to us directly by the authors and are stored as `r melnychuk_price_path`. 
> Melnychuk, M. C., Clavelle, T., Owashi, B., & Strauss, K. (2017). Reconstruction of global ex-vessel prices of fished species. ICES Journal of Marine Science, 74(1), 121–133. https://doi.org/10.1093/icesjms/fsw169

The RAM Legacy Stock Assessment Database is a compolation of stock assessments for commercially exploited species from around the world. The latest version of this database (version 4.491) was downloaded directly from the RAM website on July 25, 2020 and are stored as `r ram_sadb_path`. 
> RAM Legacy Stock Assessment Database. (2020). Version 4.491-assessment only. Released 2020-02-20. https://doi.org/10.5281/zenodo.3676087

Estimates of stock status for comericlaly exploited species not included in the RAM Legacy Stock Assessment Database from Costello et al. (2016). These data were provided to us directly by the authors and are stored as `r costello_upsides_path`. 
> Costello, C., Ovando, D., Clavelle, T., Strauss, C. K., Hilborn, R., Melnychuk, M. C., Branch, T. A., Gaines, S. D., Szuwalski, C. S., Cabral, R. B., Rader, D. N., & Leland, A. (2016). Global fishery prospects under contrasting management regimes. Proceedings of the National Academy of Sciences, 113(18), 5125–5129. https://doi.org/10.1073/pnas.1520420113

Fisheries management indicators for many of the top fishing nations from Melnychuk et al. (2017). These data were manually transcribed from Figure S2 in the supplementary information and are stored as `r melnychuk_fmi_path`. 
> Melnychuk, M. C., Peterson, E., Elliott, M., & Hilborn, R. (2017). Fisheries management impacts on target species status. Proceedings of the National Academy of Sciences, 114(1), 178–183. https://doi.org/10.1073/pnas.1609915114

The IUU Consolidated Vessel List created and maintained by Trygg Mat Tracking. These data were downloaded directly from the website on November 30, 2019 and are stored as `r iuu_vessel_path`. 
> Trygg Mat Tracking. (2020). Combined IUU Fishing Vessel List. https://www.iuu-vessels.org/

We also use the following tables to aid in our analysis:

- Proposed tier assignments for a cap-based approach to limiting subsidies as included in the proposal made to the WTO by the US Delegation on May 3, 2019 [TN/RL/GEN/197/Rev.1].

- A manually compiled table of overseas dependencies of WTO Member and Observer countries

- A manually compiled table of FAO statistical regions coorepsonding to the stocks contained in version 4.491 of the RAM Legacy Stock Assessment Database.

# Note: Country naming, political entities, and dependencies

The naming of flag and coastal states varies widely across our data sources and we recognize that this can be a politially sensitive matter. For the purposes of merging multiple data sources, we therefore translate all state names from the raw data sources to ISO-3 character codes. Country names are only used for display purposes, and the use of any particular name is not meant to convey any options regarding sovereignty on behalf of the authors.

```{r}
# Load country dependencies helper file
country_dependencies <- read_csv(country_lookup_path)

# EU countries
eu_countries <- country_dependencies$iso3[country_dependencies$is_EU]
eu_country_names <- country_dependencies$WTO_name[country_dependencies$iso3 %in% eu_countries]

# EU overseas territories
eu_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 %in% eu_countries]
eu_territory_names <- sort(c(countrycode(eu_territories, "iso3c", "country.name"), "Ascension Island", "Tristan da Cunha"))

# US overseas territories:  American Samoa, Guam, Northern Marianas Islands, Puerto Rico, US Virgin Islands
us_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "USA"]
us_territory_names <- sort(countrycode(us_territories, "iso3c", "country.name"))

# Norwegian overseas territories: 
nor_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "NOR"]
nor_territory_names <- sort(countrycode(nor_territories, "iso3c", "country.name"))
```

Additionally, it is important to note that the European Union as a political entity is a Member of the World Trade Organization, rather then the individual states that make up the European Union. We therefore need to create a profile for the European Union as a whole.  The following `r length(eu_country_names)` states are considered to be part of the European Union for the purposes of this analysis: `r paste(eu_country_names, collapse = ", ")`.

The Netherlands, France, Denmark, and the United Kingdom also have overseas dependencies that should be considered to be a part of the European Union for statistical purposes. The following overseas dependencies are also considered to be a part of the European Union for the purpose of aggregating data in this analysis: `r paste(eu_territory_names, collapse = ", ")`.

Outside of the EU, the other dependencies we need to worry about are those that relate to the United States and Norway. The following overseas dependencies are considered to be a part of the United States for the purpose of aggregating data in this analysis: `r paste(us_territory_names, collapse = ", ")`. The following overseas dependencies are considered to be a part of Norway for the purpose of aggregating data in this analysis: `r paste(norway_territory_names, collapse = ", ")`.

# Section 1 - Subsidies Data

## Updated estimates and analysis of global fisheries subsidies from Sumaila et al. (2019)

```{r}
### Data Wrangling
# Load and wrangle 2019 data
sumaila_2019_dat <- read_xlsx(sumaila_subsidies_path) %>%
  clean_names(case = "snake") %>%
  mutate(iso3 = case_when(country == "Dominican Rp" ~ "DOM",
                          country == "Micronesia" ~ "FSM",
                          country == "Untd Arab Em" ~ "ARE",
                          TRUE ~ countrycode(country, "country.name", "iso3c"))) %>%
  dplyr::select(iso3, type = class, data_type, value = constant_2018_usd) %>%
  mutate(year = 2018,
         units = "2018 US$",
         source = "Sumaila et al. (2019)") %>%
  arrange(iso3)

# Load lookup table for classification scheme
sumaila_lookup <- read_csv(here::here("data", "lookup-tables", "sumaila_subsidies_classification_lookup_table.csv"))

# Clean up the naming/spelling of the classification scheme as well as data type
sumaila_2019_dat_edit <- sumaila_2019_dat %>%
  left_join(sumaila_lookup, by = "type") %>%
  dplyr::select(iso3, year, category, category_name, type, type_name, value, data_type, units, source)

sumaila_2019_dat_edit$data_type[sumaila_2019_dat_edit$data_type == "not found evidence of subsidy"] <- "Found no evidence of subsidy"

# Deal with the EU
sumaila_2019_dat_eu <- sumaila_2019_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU",
         data_type = "Aggregate") %>%
  group_by(entity, year, category, category_name, type, type_name, data_type, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity)

# Add the EU back in
sumaila_2019_dat_edit <- sumaila_2019_dat_edit %>%
  bind_rows(sumaila_2019_dat_eu) %>%
  mutate(variable = "subsidies_Sumaila")

# Save
write_csv(sumaila_2019_dat_edit, paste0(data_dir, "sumaila_et_al_2019_subsidies_tidy.csv"))
write_csv(sumaila_2019_dat_edit, paste0(app_data_dir, "sumaila_et_al_2019_subsidies_tidy.csv"))

```

## Estimates of fisheries support provision by state from the OECD (2008 - 2018)

```{r}
# Load and wrangle OECD data (2009 - 2017)
oecd_fse_dat <- read_csv(oecd_fse_path) %>%
  clean_names(case = "snake") %>%
  dplyr::filter(measure == "USD", # only pull estimates in USD for consistancy
                country != "OECD") %>% # remove totals
  dplyr::select(iso3 = country, code = variable, year, value)

# Load OECD lookup table
oecd_fse_lookup <- read_csv(here::here("data", "lookup-tables", "oecd_fse_classification_lookup_table.csv")) %>%
  distinct(category, category_name, type, type_code, type_name)

# Match based on type - that should be plenty for the figure on SubsidyExplorer
oecd_fse_dat_edit <- oecd_fse_dat %>%
  inner_join(oecd_fse_lookup, by = c("code" = "type_code")) %>%
  mutate(data_type = "Reported",
         units = paste0(year, " US$"),
         source = "OECD FSE database") %>%
  dplyr::select(iso3, year, category, category_name, type, type_name, data_type, units, source, value)

# Deal with the EU
oecd_fse_dat_eu <- oecd_fse_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU",
         data_type = "Aggregate") %>%
  group_by(entity, year, category, category_name, type, type_name, data_type, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, type)

# Add EU back in 
oecd_fse_dat_edit <- oecd_fse_dat_edit %>%
  bind_rows(oecd_fse_dat_eu) %>%
  mutate(variable = "subsidies_OECD")

# Save
write_csv(oecd_fse_dat_edit, paste0(data_dir, "oecd_2019_fse_tidy.csv"))
write_csv(oecd_fse_dat_edit, paste0(app_data_dir, "oecd_2019_fse_tidy.csv"))

```

# Section 2 - Demographic Data

## Demographic statistics by state from the World Bank's World Development Indicators Database (2000 - 2018)

```{r}
# Load data using WDI package
wb_demo_dat <- WDI(country = "all",
                   indicator = c("gdp" = "NY.GDP.MKTP.CD",
                                 "gdp_ffa" = "NV.AGR.TOTL.CD",
                                 "population" = "SP.POP.TOTL"),
                                  start = 2000,
                                  end = 2018,
                                  extra = TRUE) %>%
  dplyr::filter(region != "Aggregates") %>%
  dplyr::select(iso3 = iso3c,
                year,
                gdp,
                gdp_ffa,
                population) %>%
  mutate(iso3 = as.character(iso3)) %>%
  arrange(iso3, year) %>%
  gather(variable, value, c("gdp", "gdp_ffa", "population")) %>%
  dplyr::filter(!is.na(value)) %>%
  mutate(units = case_when(variable == "population" ~ "individuals",
                           TRUE ~ paste0(year, " US$")),
         source = "World Bank WDI database")

# Deal with the EU + territories
wb_dat_eu <- wb_demo_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

# Deal with the US + territories
wb_dat_us <- wb_demo_dat %>%
  dplyr::filter(iso3 %in% c(us_territories, "USA")) %>%
  mutate(entity = "USA") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

# Add back in
wb_dat_edit <- wb_demo_dat %>%
  dplyr::filter(iso3 != "USA") %>%
  bind_rows(wb_dat_eu) %>%
  bind_rows(wb_dat_us)

# Export clean
write_csv(wb_dat_edit, paste0(data_dir, "world_bank_2020_wdi_tidy.csv"))
write_csv(wb_dat_edit, paste0(app_data_dir, "world_bank_2020_wdi_tidy.csv"))
```

## Total number of fishers by state from the FAO Yearbook of Fishery and Aquaculture Statistics (misc. years between 1995 and 2016)

```{r}
# Pull data and wrangle
fao_fishers_dat <- read_csv(fao_fishers_path) %>%
  mutate(iso3 = countrycode(country, "country.name", "iso3c")) %>%
  dplyr::select(-country) %>%
  gather(year, value, c('1995', '2000', '2005', '2010', '2011', '2012', '2013', '2014', '2015', '2016')) %>%
  mutate(year = as.numeric(year)) %>%
  dplyr::filter(!is.na(value)) %>%
  mutate(variable = "fishers",
         units = "individuals",
         source = "FAO Yearbook of Fishery and Aquaculture Statistics (2017)")

# Deal with EU
fao_fishers_dat_eu <- fao_fishers_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

# Add EU back in 
fao_fishers_dat_edit <- fao_fishers_dat %>%
  bind_rows(fao_fishers_dat_eu) %>%
  arrange(year, iso3)

# Export clean
write_csv(fao_fishers_dat_edit, paste0(data_dir, "fao_2017_yearbook_fishers_tidy.csv"))
write_csv(fao_fishers_dat_edit, paste0(app_data_dir, "fao_2017_yearbook_fishers_tidy.csv"))
```

## Estimated number of full-time jobs in fisheries from Teh and Sumaila (2013) in 2003

```{r}
# Pull data and wrangle
teh_fte_dat <- read_xlsx(teh_fte_path) %>%
  clean_names(case = "snake") %>%
  mutate(iso3 = case_when(country == "Dominican Rp" ~ "DOM",
                          country == "Micronesia" ~ "FSM",
                          country == "Untd Arab Em" ~ "ARE",
                          TRUE ~ countrycode(country, "country.name", "iso3c"))) %>%
  dplyr::select(-country) %>%
  mutate(variable = "fishers_fte",
         year = 2003,
         units = "full-time equivalent jobs",
         source = "Teh and Sumaila (2013)") %>%
  dplyr::select(iso3, year, variable, value = total_fte, units, source)

# Deal with EU
teh_fte_dat_eu <- teh_fte_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

# Add EU back in 
teh_fte_dat_edit <- teh_fte_dat %>%
  bind_rows(teh_fte_dat_eu)

# Export clean
write_csv(teh_fte_dat_edit, paste0(data_dir, "teh_and_sumaila_2013_fte_tidy.csv"))
write_csv(teh_fte_dat_edit, paste0(app_data_dir, "teh_and_sumaila_2013_fte_tidy.csv"))
```

# Section 3 - Fisheries Data

## Fisheries landings for 2000-2017 from the FAO Global Capture Production database (2019)

```{r}
# FAO stores a lot of characteristics in lookup tables that come with the production dataset. We need to use these tables to match country names, species names, and source
fao_country_lookup <- read_csv(paste0(fao_production_path, "CL_FI_COUNTRY_GROUPS.csv")) %>%
  clean_names(case = "snake") %>%
  dplyr::select(country = un_code,
                iso3 = iso3_code,
                name = name_en) %>%
  mutate(iso3_new = case_when(name == "Belgium-Luxembourg" ~ "BEL",
                              name %in% c("Canton and Enderbury Is", "Czechoslovakia", "Neutral Zone", "Pacific Islands Trust Tr", "Serbia and Montenegro", "Un. Sov. Soc. Rep.", "Yemen Arab Republic", "Yemen, Democratic", "Yugoslavia SFR") ~ "XXX",
                              name %in% c("Channel Islands", "Sark") ~ "GGY",
                              name == "Eswatini" ~ "SWZ",
                              name == "European Union" ~ "EU",
                              name %in% c("Johnston Island", "Midway Islands", "Wake Island", "US Minor Outlying Is.") ~ "UMI",
                              name == "Netherlands Antilles" ~ "ANT",
                              name == "Saint-Martin" ~ "MAF",
                              name == "Zanzibar" ~ "TZA",
                              name == "Other nei" ~ "UNK",
                              is.na(name) ~ "UNK",
                              TRUE ~ countrycode(name, "country.name", "iso3c"))) %>%
  dplyr::filter(iso3_new != "XXX") %>% # remove entities that no longer exist today
  dplyr::select(iso3 = iso3_new, country) %>%
  mutate(country = as.integer(country))

# FAO species lookup 
fao_species_lookup <- read_csv(paste0(fao_production_path, "CL_FI_SPECIES_GROUPS.csv")) %>%
  clean_names(case = "snake") %>%
  dplyr::select(species = x3alpha_code,
                asfis_species = name_en,
                scientific_name,
                yearbook_group,
                isscaap_group)

# FAO Symbol lookup
fao_symbol_lookup <- read_csv(paste0(fao_production_path, "CL_FI_SYMBOL.csv"))

# FAO Area lookup
fao_area_lookup <- read_csv(paste0(fao_production_path, "CL_FI_WATERAREA_GROUPS.csv")) %>%
  clean_names(case = "snake") %>%
  dplyr::select(area = code) %>%
  mutate(area = as.integer(area))

# Load data
fao_production_dat <- read.csv(paste0(fao_production_path, "TS_FI_CAPTURE.csv")) %>%
  clean_names(case = "snake") %>%
  dplyr::filter(year >= 2000) %>%
  dplyr::filter(unit == "t") %>% # remove catches measured in number of individuals
  mutate(species = as.character(species)) %>%
  left_join(fao_country_lookup, by = "country") %>%
  inner_join(fao_area_lookup, by = c("fishing_area" = "area")) %>%
  group_by(iso3, year, species, fishing_area) %>%
  summarize(value = sum(quantity, na.rm = T)) %>%
  ungroup() %>%
  left_join(fao_species_lookup, by = "species")

# Deal with the EU + territories
fao_production_dat_eu <- fao_production_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories, "EU")) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity)

# Deal with the US + territories
fao_production_dat_us <- fao_production_dat %>%
  dplyr::filter(iso3 %in% c(us_territories, "USA")) %>%
  mutate(entity = "USA-T") %>%
  group_by(entity, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity)

# Put them back together
fao_production_dat_edit <- fao_production_dat %>%
  dplyr::filter(iso3 != "EU") %>%
  bind_rows(fao_production_dat_eu) %>%
  bind_rows(fao_production_dat_us) %>%
  mutate(variable = "capture_production",
         units = "tonnes",
         source = "FAO (2019)")

write_csv(fao_production_dat_edit, paste0(data_dir, "fao_2019_capture_production_tidy.csv"))

# Aggregate by ISCCAP groups and export for time series plot
fao_production_dat_edit_groups <- fao_production_dat_edit %>%
  group_by(iso3, year, isscaap_group, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T))

write_csv(fao_production_dat_edit_groups, paste0(data_dir, "fao_2019_capture_production_isscaap_groups_tidy.csv"))
write_csv(fao_production_dat_edit_groups, paste0(app_data_dir, "fao_2019_capture_production_isscaap_groups_tidy.csv"))
```

## Ex-vessel prices from Melnychuk et al. (2017)
```{r}
# Price data from Melynchuk et al. (2017)
melnychuk_price_dat <- read_csv(melnychuk_price_path) %>%
  clean_names(case = "snake") %>%
  dplyr::select(-entry) %>%
  dplyr::filter(year >= 2000 & year <= 2012) %>%
  group_by(year, scientific_name, asfis_species, isscaap_group, group_for_pairing) %>%
  summarize(exvessel = mean(exvessel, na.rm = T)) %>%
  ungroup()

# Pull data from 2012 to use for extrapolation to 2017
melnychuk_price_dat_mry <- melnychuk_price_dat %>%
  dplyr::filter(year == 2012) %>%
  rename(exvessel_for_extrap = exvessel) %>%
  dplyr::select(-year)

# CPI for 2012 - 2018 to convert USD
cpi_2012 <- 229.594
cpi_2013 <- 232.957
cpi_2014 <- 236.736
cpi_2015 <- 237.017
cpi_2016 <- 240.007
cpi_2017 <- 245.120

# Do extrapolation for missing years
melnychuk_price_dat_extrapolation <- melnychuk_price_dat %>%
  complete(nesting(scientific_name, asfis_species, isscaap_group, group_for_pairing), year = 2000:2017) %>%
  left_join(melnychuk_price_dat_mry, by = c("scientific_name", "asfis_species", "isscaap_group", "group_for_pairing")) %>%
  mutate(exvessel_adjusted = case_when(year <= 2012 ~ exvessel,
                                       year == 2013 ~ exvessel_for_extrap * (cpi_2013/cpi_2012),
                                       year == 2014 ~ exvessel_for_extrap * (cpi_2014/cpi_2012),
                                       year == 2015 ~ exvessel_for_extrap * (cpi_2015/cpi_2012),
                                       year == 2016 ~ exvessel_for_extrap * (cpi_2016/cpi_2012),
                                       year == 2017 ~ exvessel_for_extrap * (cpi_2017/cpi_2012))) %>%
  dplyr::select(-exvessel, -exvessel_for_extrap) %>%
  rename(exvessel = exvessel_adjusted)

# Get rid of NAs and save
melnychuk_price_dat_extrapolation <- melnychuk_price_dat_extrapolation %>%
  dplyr::filter(!is.nan(exvessel)) %>%
  mutate(variable = "ex_vessel_price",
         units = paste(year, "US$"),
         source = "Melnychuk et al. (2017)")

write_csv(melnychuk_price_dat_extrapolation, paste0(data_dir, "melnychuk_2017_price_tidy.csv"))

# Create averages by pairing group for each year
price_dat_group_year <- melnychuk_price_dat_extrapolation %>%
  group_by(year, group_for_pairing) %>%
  summarize(exvessel_group_avg = mean(exvessel, na.rm = T)) %>%
  ungroup()

# Create averages by pairing group (no year) 
price_dat_group <- melnychuk_price_dat_extrapolation %>%
  group_by(group_for_pairing) %>%
  summarize(exvessel_group_avg_allyears = mean(exvessel, na.rm = T)) %>%
  ungroup()
```

## Estimated landed value from FAO/Melnychuk et al. (2017)

```{r}
# Combine landings and price data
combine_catch_price_dat <- fao_production_dat_edit %>%
  left_join(melnychuk_price_dat_extrapolation %>% dplyr::select(-isscaap_group, -variable, -source, -units), by = c("year", "scientific_name", "asfis_species", "isscaap_group" = "group_for_pairing")) %>%
  left_join(price_dat_group_year, by = c("year", "isscaap_group" = "group_for_pairing")) %>%
  left_join(price_dat_group, by = c("isscaap_group" = "group_for_pairing")) %>%
  mutate(exvessel_inclusive = case_when(!is.na(exvessel) ~ exvessel,
                                        !is.na(exvessel_group_avg) ~ exvessel_group_avg,
                                        !is.na(exvessel_group_avg_allyears) ~ exvessel_group_avg_allyears))

# Look at NAs - 9398 entries
combine_nas <- combine_catch_price_dat %>%
  dplyr::filter(is.na(exvessel_inclusive))

# Set price to 0 for those
combine_catch_price_dat$exvessel_inclusive[is.na(combine_catch_price_dat$exvessel_inclusive)] <- 0

# Calculate revenues and export
landed_value_dat <- combine_catch_price_dat %>%
  dplyr::select(-exvessel, -exvessel_group_avg, -exvessel_group_avg_allyears) %>%
  rename(exvessel = exvessel_inclusive) %>%
  mutate(landed_value = value*exvessel) %>%
  group_by(iso3, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(landed_value, na.rm = T)) %>%
  mutate(variable = "landed_value",
         units = paste(year, "US$"),
         source = "FAO (2019); Melnychuk et al. (2017)")

write_csv(landed_value_dat, paste0(data_dir, "estimated_landed_value_tidy.csv"))

# Calculate revenues and aggregate by ISSCAAP group
landed_value_dat_groups <- landed_value_dat %>%
  group_by(iso3, year, isscaap_group, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T))

write_csv(landed_value_dat_groups, paste0(data_dir, "estimated_landed_value_isscaap_groups_tidy.csv"))
write_csv(landed_value_dat_groups, paste0(app_data_dir, "estimated_landed_value_isscaap_groups_tidy.csv"))
```

## Stock status database
```{r}
# Load data
costello_upsides_dat <- read_csv(costello_upsides_path) %>%
  clean_names(case = "snake")

# We need to split the entries that span multiple FAO regions...  
costello_upsides_dat_edit <- costello_upsides_dat %>%
  mutate(region_fao_chr = as.character(region_fao),
         n_regions = nchar(region_fao_chr)/2,
         region_fao_split = str_split(gsub("(.{2})", "\\1 ", region_fao_chr), " ")) %>%
  unnest(cols = c(region_fao_split)) %>%
  mutate(region_fao_split = as.numeric(region_fao_split)) %>%
  dplyr::filter(region_fao_split != "") %>%
  mutate(iso3 = case_when(country == "Channel Islands" ~ "JEY",
                          country == "Micronesia" ~ "FSM",
                          country == "Zanzibar" ~ "TZA",
                          TRUE ~ countrycode(country, "country.name", "iso3c")))

# Save clean
write_csv(costello_upsides_dat_edit, paste0(data_dir, "costello_et_al_2016_upsides_tidy.csv"))

```

# Melnychuk FMI
```{r}
# Load data and wrangle
melnychuk_fmi_dat <- read_csv(melnychuk_fmi_path) %>%
  clean_names(case = "snake") %>%
  mutate(iso3 = countrycode(country, "country.name", "iso3c")) %>%
  dplyr::select(-country) %>%
  gather(variable, value, -iso3)

# Create a mean aggregate for the EU
melnychuk_fmi_dat_eu <- melnychuk_fmi_dat  %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, variable) %>%
  summarize(value = mean(value, na.rm = T)) %>%
  rename(iso3 = entity)

# Add EU back in
melnychuk_fmi_dat_edit <- melnychuk_fmi_dat %>%
  bind_rows(melnychuk_fmi_dat_eu)

# Export clean
write_csv(melnychuk_fmi_dat_edit, paste0(data_dir, "melnychuk_et_al_2017_fmi_tidy.csv"))

```

# Section 4 - Other Data

## Cap and Tier Data from the US Proposal
```{r}
cap_tier_dat <- read_csv(cap_tier_path) %>%
  mutate(iso3 = case_when(country == "Eswatini" ~ "SWZ",
                          country == "EU" ~ "EU",
                          TRUE ~ countrycode(country, "country.name", "iso3c"))) %>%
  dplyr::select(-country)

write_csv(cap_tier_dat, paste0(data_dir, "USA_cap_tier_tidy.csv"))
write_csv(cap_tier_dat, paste0(app_data_dir, "USA_cap_tier_tidy.csv"))
```

# Section 5 - Relative Metrics

This section is a bit of a relic from when we were doing "country profiles", but there are a couple of relative metrics we want to calculate from our data sources. This is done here.

```{r}
# Sumaila dat for matching
subs_dat_match <- sumaila_2019_dat_edit %>%
  dplyr::select(-units, -variable) %>%
  rename(subs = value, subs_year = year, subs_source = source)

# GDP
gdp_dat_mry <- wb_dat_edit %>%
  dplyr::filter(variable == "gdp") %>%
  group_by(iso3, variable) %>%
  dplyr::filter(year == max(year[!is.na(value)])) %>%
  ungroup() %>%
  dplyr::select(iso3, gdp_year = year, gdp = value, gdp_source = source) %>%
  right_join(subs_dat_match, by = c("iso3")) %>%
  mutate(variable = "subsidies_per_gdp",
         data_type = paste0(data_type),
         source = paste0(subs_source, "; ", gdp_source),
         value = subs/gdp,
         units = paste0(subs_year, " US$/", gdp_year, " US$")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, variable, value, units, data_type, source) %>%
  dplyr::filter(!is.na(value))

# Population
pop_dat_mry <- wb_dat_edit %>%
  dplyr::filter(variable == "population") %>%
  group_by(iso3, variable) %>%
  dplyr::filter(year == max(year[!is.na(value)])) %>%
  ungroup() %>%
  dplyr::select(iso3, population_year = year, population = value, population_source = source) %>%
  right_join(subs_dat_match, by = c("iso3")) %>%
  mutate(variable = "subsidies_per_capita",
         data_type = paste0(data_type),
         source = paste0(subs_source, "; ", population_source),
         value = subs/population,
         units = paste0(subs_year, " US$/person (", population_year, ")")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, variable, value, units, data_type, source) %>%
  dplyr::filter(!is.na(value))

# Fishers
fisher_dat_mry <- fao_fishers_dat_edit %>%
  group_by(iso3, variable) %>%
  dplyr::filter(year == max(year[!is.na(value)])) %>%
  ungroup() %>%
  dplyr::select(iso3, fishers_year = year, fishers = value, fishers_source = source) %>%
  right_join(subs_dat_match, by = c("iso3")) %>%
  mutate(variable = "subsidies_per_fisher",
         data_type = paste0(data_type),
         source = paste0(subs_source, "; ", fishers_source),
         value = subs/fishers,
         units = paste0(subs_year, " US$/fisher (", fishers_year, ")")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, variable, value, units, data_type, source) %>%
  dplyr::filter(!is.na(value))

# FTE
fte_dat_mry <- teh_fte_dat_edit %>%
  dplyr::select(iso3, fte_year = year, fte = value, fte_source = source) %>%
  right_join(subs_dat_match, by = c("iso3")) %>%
  mutate(variable = "subsidies_per_fte",
         data_type = paste0(data_type),
         source = paste0(subs_source, "; ", fte_source),
         value = subs/fte,
         units = paste0(subs_year, " US$/full-time-equivalent (", fte_year, ")")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, variable, value, units, data_type, source) %>%
  dplyr::filter(!is.na(value))

# Capture Production
production_dat_mry <- fao_production_dat_edit_groups %>%
  group_by(iso3, year, variable) %>%
  summarize(value = sum(value, na.rm = T),
            source = unique(source)) %>%
  group_by(iso3, variable) %>%
  dplyr::filter(year == max(year[!is.na(value)])) %>%
  ungroup() %>%
  dplyr::select(iso3, production_year = year, production = value, production_source = source) %>%
  right_join(subs_dat_match, by = c("iso3")) %>%
  mutate(variable = "subsidies_per_production",
         data_type = paste0(data_type),
         source = paste0(subs_source, "; ", production_source),
         value = subs/production,
         units = paste0(subs_year, " US$/tonne (", production_year, ")")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, variable, value, units, data_type, source) %>%
  dplyr::filter(!is.na(value))

# Landed Value
landed_value_dat_mry <- landed_value_dat %>%
  group_by(iso3, year, variable) %>%
  summarize(value = sum(value, na.rm = T),
            source = unique(source)) %>%
  group_by(iso3, variable) %>%
  dplyr::filter(year == max(year[!is.na(value)])) %>%
  ungroup() %>%
  dplyr::select(iso3, landed_value_year = year, landed_value = value, landed_value_source = source) %>%
  right_join(subs_dat_match, by = c("iso3")) %>%
  mutate(variable = "subsidies_per_landed_value",
         data_type = paste0(data_type),
         source = paste0(subs_source, "; ", landed_value_source),
         value = subs/landed_value,
         units = paste0(subs_year, " US$/", landed_value_year, " US$")) %>%
  dplyr::select(iso3, category, category_name, type, type_name, variable, value, units, data_type, source) %>%
  dplyr::filter(!is.na(value))

# Combine with subsidy data and calculate relative metrics
relative_dat <-  gdp_dat_mry %>%
  bind_rows(pop_dat_mry) %>%
  bind_rows(fisher_dat_mry) %>%
  bind_rows(fte_dat_mry) %>%
  bind_rows(production_dat_mry) %>%
  bind_rows(landed_value_dat_mry)

write_csv(relative_dat, paste0(data_dir, "relative_subsidy_metrics_tidy.csv"))
write_csv(relative_dat, paste0(app_data_dir, "relative_subsidy_metrics_tidy.csv"))
```