### 2020-08-07 scripts/06_run_model.Rmd

---
output:
  html_document: default

title: "SubsidyExplorer - Step 5: Run Model"
author: "Kat Millage, Vienna Saccomanno, Laura Lea Rubino, Christopher Costello"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
bibliography: "../pew-subsidies.bib"
---

```{r setup, include=FALSE}
# Load packages
library(knitr) 
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This script generates the most ambitous global and regional results obtained by removing all capacity enhancing subsidies from all WTO Members and Observers.

```{r}
# Source model script
source(here::here("app", "scripts", "BioEconModel.R"))
```

```{r}
# Load vessel list
pro_rate_subsidies <- F

vessel_dat <- read.csv(here::here("app", "data", "vessel_list_2018_final.csv"), stringsAsFactors = F)

if(pro_rate_subsidies == T){

  vessel_dat <- vessel_dat %>%
    mutate(B1_subs = B1_subs * 0.54, # boat construction/renovation - matched to payments based on vessels
           B2_subs = B2_subs * 1, # fishery development projects/support services - matched to payments based on variable use
           B3_subs = B3_subs * 0.87, # port construction and renovation - matched to payments based on vessels
           B4_subs = B4_subs * 0.87, # price/marketing support, processing infrastructure - matched to payments based on output
           B5_subs = B5_subs * 0.76, # non-fuel tax exemptions - matched to payments based on fishers income
           B6_subs = B6_subs * 0.56, # foreign access agreements - matched to payments based on fishers own capital
           B7_subs = B7_subs * 0.84, # fuel - matched to payments based on fuel use
           bad_subs = (B1_subs + B2_subs + B3_subs + B4_subs + B5_subs + B6_subs + B7_subs))

}

vessel_dat <- vessel_dat %>%
  mutate(eez_hs_code = case_when(eez_id == 0 ~ paste0("HS-", fao_region),
                                 TRUE ~ as.character(eez_id)))
```

```{r}
# Load biological parameters for the model
bio_dat_global <- read.csv(here::here("app", "data", "model_parameters_global.csv"))
bio_dat_regional <- read.csv(here::here("app", "data", "model_parameters_regional.csv"))

# Regional parameter list
bio_dat_regional_list <- bio_dat_regional %>%
  gather(region, value, -c(1:2)) %>%
  group_by(region) %>%
  group_split()
```

```{r}
# Get WTO countries territories
```

```{r}
# Threshold to be considered managed
managed_cutoff <- 0.66
```

## Create global fleet

```{r pressure, echo=FALSE}
# Assign fleets to vessels
global_fleet_vessels <- vessel_dat %>%
  mutate(is_WTO = ifelse(flag_iso3 %in% c(cap_tier_dat$iso3, eu_countries, eu_territories, us_territories, norwegian_territories), T, F)) %>%
  mutate(fleet = case_when(is_WTO & fmi_best >= managed_cutoff & bad_subs > 0 ~ "affected_managed",
                           is_WTO & fmi_best < managed_cutoff & bad_subs > 0 ~ "affected_oa",
                           (!is_WTO & fmi_best >= managed_cutoff) ~ "unaffected_managed",
                           (!is_WTO & fmi_best < managed_cutoff) ~ "unaffected_oa",
                           TRUE ~ "unaffected_oa"))

# Create fleet (summary)
global_fleet_summary <- global_fleet_vessels %>%
  group_by(region, fleet) %>%
  summarize(catch = sum(catch, na.rm = T),
            bad_subs = sum(bad_subs, na.rm = T),
            fishing_KWh = sum(fishing_KWh_eez_fao_ter, na.rm = T),
            removed_subs = sum(bad_subs, na.rm = T))
```

## Run Model (Remove all bad subsidies)

```{r}

```